Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-03-17 Manuscript/HI_Model_ms.Rmd
updated manuscript files

>>>>

Our model does not currently differentiate between forests dominated by native versus non-native tree species, which could influence estimates of ecosystem carbon balance. Native forests in the Hawaiian Islands are dominated by ʻōhiʻa (_Metrosideros polymorpha_ Gauditch), an endemic foundational tree species found across a broad range of climatic and edaphic conditions [@ziegler_hawaiian_2002]. Beginning in 2010, a fungal disease termed Rapid ʻŌhiʻa Death (ROD) caused by two _Ceratocystus_ spp. has emerged and caused widespread mortality to mature ʻōhiʻa trees across a range of size classes, primarily on Hawaiʻi Island [@mortenson_assessing_2016; @fortini_evolving_2019]. The distribution and potential range of this emerging threat to Hawaiʻi's most widespread native tree species has only recently been mapped [@vaughn_approach_2018; @fortini_evolving_2019], but the pulse of newly dead organic matter and reduction in photosynthetic capacity induced by widespread tree mortality could significantly alter ecosystem carbon balance over the long-term [@sleeter_effects_2019]. ROD-affected forest canopies could also potentially see a replacement of long-lived, stress-tolerating ʻōhiʻa with non-native tree species that have faster growth rates, lower wood density, and faster tissue turnover, all of which could influence the long-term trajectory of forest carbon cycling in Hawaiʻi. New model projections for Hawaiʻi that incorporate ROD spread rates and forest restoration scenarios will therefore require differentiation among forest ecosystems dominated by native and non-native tree species.
